---
title: "At_PGPR_PVY"
author: "Bibechana Paudel"
date: "2025-10-29"
output: html_document
---
# **First trial**

###Libraries used

```{r}
#install.packages("tidyverse")
#install.packages("ggplot2")
#install.packages("agricolae")
#install.packages("tidyr")
#install.packages("ggpubr")
#install.packages("multcomp")
#install.packages("emmeans")
#install.packages("multcompView")
library(tidyverse)
library(ggplot2)
library(agricolae)
library(tidyr)
library(ggpubr)
library(multcomp)
library(emmeans)
library(multcompView)
```

###Load and verify data of Rep 1
```{r}
rep <- read.csv("Data-obj3-New/At_PGPR_PVY_New.csv", header=T)
head(rep)  # View the first few rows
str(rep)   # Check the structure of the data
summary(rep) # Quick statistical summary
```

###Make the factor variable
```{r}
rep$Treatment<-as.factor(rep$Treatment)
rep$Dpi<-as.factor(rep$Dpi)
```

##Calculate the ViralLoad using the regression equation derived by (Feng , J.L et al., 2006 ) and log transform for normality.

[Feng, J.L et al., 2006](https://academic.oup.com/abbs/article/38/10/669/217)

```{r}
rep<-rep %>% 
mutate (ViralLoad=((10^((Cq-49.153)/-3.93)))) %>% ##added a new column and named ViralLoad
mutate(logViralLoad=log10(ViralLoad)) 
 ##added a new column and named logViralLoad
```
## Statistical analysis
```{r}
library(dplyr)
library(agricolae)
library(purrr)
library(tibble)

# get all unique Dpi values
dpi_levels <- unique(rep$Dpi)

# function: fit lm -> run LSD.test -> extract groups
run_lsd_per_dpi <- function(dpi_label) {
  d <- subset(rep, Dpi == dpi_label)
  fit <- aov(logViralLoad ~ Treatment, data = d)
  lsd <- LSD.test(fit, "Treatment", p.adj = "none", group = TRUE)  # Fisher's LSD
  
  # Extract the group letters
  out <- lsd$groups
  out$Treatment <- rownames(out)
  rownames(out) <- NULL
  out <- as_tibble(out) %>% mutate(Dpi = dpi_label, .before = 1)
  return(out)
}

# run for all DPIs
lsd_results <- map_dfr(dpi_levels, run_lsd_per_dpi)

# inspect
lsd_results

```


##Data summarization

```{r}
PVY1 <- rep %>%  # Start with the dataset `rep` and pipe (%>%) it into the next operations
  group_by(Treatment, Dpi) %>%   # Group the data by treatment type and days post-inoculation (Dpi)
  dplyr::summarize(
    avg.virus= mean(logViralLoad, na.rm=TRUE),
    n=n(),
    se = sd(logViralLoad)/sqrt(n)) %>% 
  left_join(lsd_results,by=c("Dpi", "Treatment"))# join the significant letters with the PVY1 data frame.
```

##Classify the sample collection days based on leaf

```{r}
PVY1$Leaf_sample <- ifelse(
  grepl("^IL", PVY1$Dpi), "Inoculated",  # Matches Dpi values starting with IL
  ifelse(grepl("^SL", PVY1$Dpi), "Systemic", NA)
)
```

## Visualization: Inoculated sample

```{r}
inoculated<-filter(PVY1, Dpi=="IL-1" | Dpi=="IL-4" | Dpi=="IL-7" | Dpi=="IL-10") #Subset the PVY1 data to include only specific timepoints labeled "IL-1", "IL-4", "IL-7", and "IL-10".
inoculated$Treatment<-factor(inoculated$Treatment, levels=c("Control","P. fluorescens", "S. marcescens", "B. subtilis"))
```


```{r PVY on IL_1st trial_At }
offset <- 0.05 * diff(range(inoculated$avg.virus + inoculated$se, na.rm = TRUE))
cbbPalette <- c( "#56B4E9","#F0E442","#CC79A7", "#E69F00")  ##Load colour blind friendly cbbPalette
PVY_IL<-ggplot(inoculated, aes(x =factor(Dpi,levels=c("IL-1", "IL-4", "IL-7", "IL-10")), y =avg.virus , fill = Treatment)) +
  stat_summary(fun=mean,geom="bar", position = position_dodge(width = 0.7), width = 0.6) +
   geom_errorbar(aes(ymin = avg.virus - se, 
                    ymax = avg.virus + se),
                width = 0.4,
                position = position_dodge(width = 0.7)) +
  geom_text(aes(y = avg.virus + se + offset, label = groups),position = position_dodge(width = 0.7), vjust = -0.5,size=4.5)+
  theme_classic() +  #add the extracted significance letter to the bar
  labs(title= "" ,
       x = "",
       y = "Log Copy Number") +  # labels of x and y axis
  scale_fill_manual(values=cbbPalette,
                    labels=c(expression(bold("Control")),expression(bolditalic("P. fluorescens")),expression(bolditalic("S. marcescens")),expression(bolditalic("B. subtilis"))))+
  scale_y_continuous(
    expand = expansion(mult = c(0, 0.1)), breaks = seq(1, 9, by = 1)  # This removes padding at the bottom of the plot
  )+
  theme(legend.position = "right",
        legend.key.size = unit(0.5, "cm"),  # Adjusts the size of the legend keys
        legend.title = element_blank(),  # Adjusts the legend title font size
        legend.text = element_text(size = 11),
               axis.text.x = element_text(color = "black", face="bold", size=11),   # X-axis text color
        axis.text.y = element_text(color = "black", face="bold", size=11),   # Y-axis text color
        axis.title.x = element_text(color = "black"),  # X-axis label color
        axis.title.y = element_text(color = "black", face="bold", size=11),  # Y-axis label color
        axis.line = element_line(color = "black")      # Axis lines color
  )

 PVY_IL
```

## Visualization: Systemic sample

```{r}
systemic<-filter(PVY1, Dpi=="SL-7" | Dpi=="SL-10" | Dpi=="SL-14")
systemic$Treatment<-factor(systemic$Treatment, levels=c("Control","P. fluorescens", "S. marcescens",  "B. subtilis"))
```


```{r,PVY on SL_1st trial_At }
offset <- 0.05 * diff(range(systemic$avg.virus + systemic$se, na.rm = TRUE))
cbbPalette <- c( "#56B4E9","#F0E442","#CC79A7", "#E69F00")
PVY_SL<-ggplot(systemic, aes(x =factor(Dpi,levels=c("SL-7","SL-10","SL-14")), y =avg.virus , fill = Treatment)) +
  stat_summary(fun=mean,geom="bar", position = position_dodge(width = 0.7), width = 0.6) +
   geom_errorbar(aes(ymin = avg.virus - se, 
                    ymax = avg.virus + se),
                width = 0.4,
                position = position_dodge(width = 0.7)) +
geom_text(aes(y = avg.virus + se + offset, label = groups),position = position_dodge(width = 0.7), vjust = -0.5, size=4.5)+   ## Add group comparison letters
  theme_classic() +
  labs(title= "" ,
       x = "",
       y = "Log Copy Number") +
  scale_fill_manual(values=cbbPalette,
                    labels=c(expression(bold("Control")),expression(bolditalic("P. fluorescens")),expression(bolditalic("S. marcescens")),expression(bolditalic("B. subtilis"))))+
  scale_y_continuous(
    expand = expansion(mult = c(0, 0.1)), breaks = seq(1, 9, by = 1)  # This removes padding at the bottom of the plot
  )+
  theme(legend.position = "right",
        legend.key.size = unit(0.5, "cm"),  # Adjusts the size of the legend keys
        legend.title = element_blank(),  # Adjusts the legend title font size
        legend.text = element_text(size = 11),
                axis.text.x = element_text(color = "black", face="bold", size=11),   # X-axis text color
        axis.text.y = element_text(color = "black", face="bold", size=11),   # Y-axis text color
        axis.title.x = element_text(color = "black"),  # X-axis label color
        axis.title.y = element_text(color = "black", face="bold", size=11),  # Y-axis label color
        axis.line = element_line(color = "black")      # Axis lines color
  )

 PVY_SL
```

## Combined figure

```{r, Combine fig 1st trial}
PVY<-ggarrange(PVY_IL, PVY_SL,    #combine the  figures 
                   nrow=1,
                   ncol=2,
                   common.legend=T,    # Gives the common legend for both figures
                   widths=c(1.25,1))   # Makes the width of 1st fig 1.75 times more than 2nd fig
PVY
```


